---
title: "Functional role of SMPD1 in PDAC"
subtitle: "__graph pad__"
author: "_umahajan_"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_notebook:
    theme: united
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: false
---

```{r setup, include=FALSE}
chooseCRANmirror(graphics=TRUE, ind=1)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=85),tidy=TRUE, echo=TRUE, warning=FALSE, message=FALSE)
```

# load packages and datasets

```{r packages}
rm(list = ls())
##----------------------------------------------------------------
##                        load packages                        --
##----------------------------------------------------------------
scriptLibraries <-  c(
  "here",
  "tidyverse",
  "readxl",
  "RColorBrewer",
  "ggpubr",
  "rstatix",
  "drc",
  "scales",
  "FSA"
  
)

#devtools::install_github("umahajanatlmu/metapacR", ref="master", auth_token = "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")

##---------------------------------------------------------------
##                      load functions                         --
##---------------------------------------------------------------
devtools::source_url("https://raw.githubusercontent.com/umahajanatlmu/useful_commands/main/auxillary/basicFunctions.R")

# load packages 
##---------------------------------------------------------------
installScriptLibs(scriptLibraries)

## results directory
ifelse(!dir.exists(file.path(paste0(here()), "results_graphpad")),
       dir.create(file.path(paste0(here()), "results_graphpad")), FALSE)

save_plot <- function(x, filename, width = 7, height = 7) {
    stopifnot(!missing(x))
    stopifnot(!missing(filename))
    svg(filename, width = width, height = height)
    grid::grid.newpage()
    grid::grid.draw(x)
    dev.off()
}
```

# plot graphs

## Oroboros
```{r}
tables <- pzfx::pzfx_tables("20260131_Oroboros/Results_Wt_vs_C5/Oroborus_Analysis_WTvsC5.pzfx")

## create dataframe

oro_df <- data.frame()

i <- 1
for (t in tables) {
  
  print(t)
  df <- pzfx::read_pzfx("20260131_Oroboros/Results_Wt_vs_C5/Oroborus_Analysis_WTvsC5.pzfx", table = t) %>%
    pivot_longer(cols = everything(),names_to = "Clone", values_to = "conc") %>%
    mutate(conc = gsub(",",".", conc)) %>%
    mutate(resp = t, index = i)
  
  oro_df <- bind_rows(oro_df,df)
  
  i <- i +1
  
}

oro_df$conc <- as.numeric(oro_df$conc)
oro_df$clone <- ifelse(oro_df$Clone == "WT", "WT", "KO")
oro_df$clone <- factor(oro_df$clone, levels=c("WT", "KO"))



# Order the factor levels by your index column
oro_df$resp <- factor(oro_df$resp, levels = unique(oro_df$resp[order(oro_df$index)]))

p <- ggplot(data=oro_df) +
  geom_boxplot(aes(x= clone, 
                  y= conc,
                  fill=clone), 
               color="black",
               show.legend = FALSE) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  ggnewscale::new_scale_fill() +
  geom_jitter(aes(x= clone, 
                  y= conc, 
                  fill=Clone),
              size=3, 
              width = 0.4, 
              shape = 21, 
              color= "black",
              alpha= 0.8,
               show.legend = FALSE) +
  scale_fill_manual(values = c("#EF8A62","#D1E5F0")) +
  theme_publication() +
  xlab("") +
  ylab("Specific O2 flux (pmol/s*ml)") +
  facet_wrap(~resp, scales = "free_y") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  stat_compare_means(data=oro_df, 
                     aes(x= clone, 
                         y= conc), 
                     label = "..p..",
                     method = "t.test",
                     show.legend = FALSE)


print(p)

ggsave("oroboros.pdf",
       plot = p, 
       width = 18,
       height = 15,
       dpi = 300,
       units = "cm") 

```

## fcs data
```{r}
library(flowCore)
library(flowViz)
library(ggplot2)

# Your data is already loaded as 'fs'
fs <- read.flowSet(path = "20260101_FCS_Cell_Cycle/", pattern = ".fcs", transformation = FALSE)

# Check if all files have the same channels
fsApply(fs, function(ff) parameters(ff)$name)

# 1. Quality Control - Gate on singlets using FSC-A vs FSC-H
# Create a singlet gate (adjust coordinates based on your data)
singlet_gate <- rectangleGate(filterId="Singlets",
                               "FSC-A"=c(0, 250000),
                               "FSC-H"=c(0, 250000))

# Apply singlet gating
fs_singlets <- Subset(fs, singlet_gate)

# 2. Gate on viable cells using FSC-A vs SSC-A
viable_gate <- rectangleGate(filterId="Viable",
                              "FSC-A"=c(50000, 250000),
                              "SSC-A"=c(0, 250000))

fs_viable <- Subset(fs_singlets, viable_gate)

# 3. Extract PE-A (DNA content) data
# View histogram of negative control first
neg_control <- fs_viable[["NG_Tube_001_013.fcs"]]
plot(density(exprs(neg_control)[,"PE-A"]), 
     main="Negative Control - PE-A Distribution",
     xlab="PE-A (DNA Content)")

# 4. Cell cycle gating (adjust thresholds based on your data)
# You'll need to identify G0/G1, S, and G2/M peaks
# Example gates (ADJUST THESE based on your histogram):
G1_gate <- rectangleGate(filterId="G1", "PE-A"=c(2000, 4000))
S_gate <- rectangleGate(filterId="S", "PE-A"=c(4000, 5000))
G2M_gate <- rectangleGate(filterId="G2M", "PE-A"=c(5000, 50000))

# 5. Calculate cell cycle percentages for each sample
calculate_cell_cycle <- function(ff) {
  pe_data <- exprs(ff)[,"PE-A"]
  
  
  # Adjust these thresholds based on your data
  G1_count <- sum(pe_data > 2000 & pe_data < 4000)
  S_count <- sum(pe_data > 4000 & pe_data <5000)
  G2M_count <- sum(pe_data > 5000 & pe_data < 50000)
  
  total <- G1_count + S_count + G2M_count
  
  return(c(G1=G1_count/total*100, 
           S=S_count/total*100, 
           G2M=G2M_count/total*100))
}

# Apply to all samples (excluding negative control)
fs_samples <- fs_viable[sampleNames(fs_viable) != "NG_Tube_001_013.fcs"]
cell_cycle_results <- fsApply(fs_samples, calculate_cell_cycle)

# 6. Create results dataframe
results_df <- as.data.frame(cell_cycle_results)
results_df$Sample <- rownames(results_df)
results_df$clone_id <- gsub("_R[0-9].*", "", results_df$Sample)

print(results_df)

# 7. Visualize results
library(reshape2)
results_long <- melt(results_df, id.vars=c("Sample", "clone_id"))  
results_long$clone <- ifelse(results_long$clone_id == "WT", "WT", "KO")

results_long$clone_id <- factor(results_long$clone_id, c("WT", "C3", "C5", "C6"))

results_long$clone <- factor(results_long$clone, c("WT", "KO"))

plot <- ggplot(results_long) +
  stat_summary(aes(x = variable, y = value, fill = clone),
               fun = mean, geom = "bar",
               position = position_dodge(0.9), color = "black") +
  stat_summary(aes(x = variable, y = value, fill = clone),
               fun.data = mean_sd, geom = "errorbar",
               width = 0.5, position = position_dodge(0.9)) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  ggnewscale::new_scale_fill() +
  geom_point(aes(x = variable, 
                 y = value, 
                 fill = clone_id,
                 group = clone),  
             size = 3, 
             position = position_jitterdodge(jitter.width = 0.4, 
                                             dodge.width = 0.9),
             shape = 21, 
             color = "black",
             alpha = 0.8) +
  stat_compare_means(aes(x = variable, 
                         y = value,
                         group = clone), 
                     method = "t.test",
                     method.args = list(formula = value ~ variable),
                     label = "p.format", 
                     na.rm = TRUE, bracket.size = 1) +
  scale_fill_manual(values = c("#D1E5F0","#B2182B","#EF8A62","#FDDBC7")) +
  xlab("") +
  ylab("Percentage (%)") +
  theme_publication()

plot

ggsave("cell_cycle.pdf",
       plot = plot, 
       width = 9,
       height = 8,
       dpi = 300,
       units = "cm")
```

## lysotracker
```{r}
dat <- read.csv("20260202_lysotracker/output/LysT_lysotracterObjects.csv") %>%
  dplyr::select(Metadata_clone,AreaShape_Area,Intensity_MeanIntensity_lysotracker)

colnames(dat) <- c("Clone_ID", "Area", "Intensity")

dat <- reshape2::melt(dat, id.vars=c("Clone_ID"))

dat$Clone_ID <- toupper(dat$Clone_ID)

dat$Clone <- ifelse(dat$Clone_ID == "WT", "WT", "KO")

dat$Clone <- factor(dat$Clone, c("WT", "KO"))

dat <- dat[dat$variable == "Area",]

p <- ggplot(data=dat) +
  geom_violin(aes(x= Clone, 
                  y= value,
                  fill=Clone), 
               color="black",
              alpha = 0.5,
               show.legend = FALSE) +
  geom_boxplot(aes(x= Clone, 
                  y= value,
                  fill=Clone), 
               width=0.1,
               color="black",
               show.legend = FALSE) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  stat_compare_means(data=dat, 
                     aes(x= Clone, 
                         y= value), 
                     label = "..p..",
                     show.legend = FALSE) +
  theme_publication() +
  xlab("") +
  ylab("Arbitrary Units") +
  ggtitle("Lysosomes")

print(p)


ggsave("lysotracker.pdf",
       plot = p, 
       width = 6,
       height = 8,
       dpi = 300,
       units = "cm")
```


## lysotracker kras
```{r}
dat <- read.csv("../../Data/KRAS_IF/output_dist/LysT_RelateObjects_nonperinuclear_lysosomes.txt",sep = "") %>%
  dplyr::select(Metadata_clone,Metadata_replicate,AreaShape_Area,Intensity_MeanIntensity_lysotracker) 

colnames(dat) <- c("Clone_ID", "replicate","Area","Intensity")

dat <- reshape2::melt(dat, id.vars=c("Clone_ID", "replicate"))

dat <- dat %>%
  group_by(Clone_ID, replicate, variable) %>%
  summarise_if(is.numeric, median) %>%
  ungroup()

dat$Clone_ID <- toupper(dat$Clone_ID)

dat$Clone <- ifelse(dat$Clone_ID == "WT", "WT", "KO")

dat$Clone <- factor(dat$Clone, c("WT", "KO"))


dat <- dat[dat$variable == "Intensity",]

p <- ggplot(data=dat) +
  geom_boxplot(aes(x= Clone, 
                  y= value,
                  fill=Clone), 
               color="black",
               show.legend = FALSE) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  ggnewscale::new_scale_fill() +
  geom_jitter(aes(x= Clone, 
                  y= value, 
                  fill=Clone_ID),
              #size=3, 
              width = 0.4, 
              shape = 21, 
              color= "black",
              alpha= 0.8) +
  stat_compare_means(data=dat, 
                     aes(x= Clone, 
                         y= value), 
                     label = "..p..",
                     show.legend = FALSE) +
  scale_fill_manual(values = c("#B2182B","#EF8A62","#FDDBC7", "#D1E5F0")) +
  theme_publication() +
  xlab("") +
  ylab("Area (AU)") +
  ggtitle("Kras in lysosomes")

print(p)

ggsave("kras_lysosomes.pdf",
       plot = p, 
       width = 9,
       height = 10,
       dpi = 300,
       units = "cm") 

```


## orthotopic injection
```{r}
colony <- read_excel("20260205_invivo_expt/Animal data.xlsx", 1)

colony <- separate(colony, col="cells", into = c("clone", "treatment"))


colony$clone <- factor(colony$clone, levels=c("WT", "C3"))
colony$treatment <- factor(colony$treatment, levels=c("control", "treat"))

p <- ggplot(data=colony) +
  geom_boxplot(aes(x= clone, 
                  y= `Weight (Pancreas+tumor) (gm)`,
                  fill=clone), 
               color="black",
               show.legend = FALSE) +
  geom_jitter(aes(x= clone, 
                  y= `Weight (Pancreas+tumor) (gm)`,
                  fill=clone), size=3, 
              width = 0.4, 
              shape = 21, 
              color= "black",
              alpha= 0.8) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  stat_compare_means(data=colony, 
                     aes(x= clone, 
                         y= `Weight (Pancreas+tumor) (gm)`), 
                     label = "..p..",
                     show.legend = FALSE) +
  theme_publication() +
  xlab("") +
  ylab("Tumor weight in g") +
  ggtitle("Tumor weight") +
  facet_wrap(~treatment)

print(p)

ggsave("20260205_invivo_expt/tumor_weights_mrtx.pdf",
       plot = p, 
       width = 10,
       height = 10,
       dpi = 300,
       units = "cm")

## spinder plot
df <- read_excel("20260205_invivo_expt/Animal data.xlsx", 2) %>%
  separate(col = "cells", into = c("clone", "treatment")) %>%
  mutate(across(-c(Animal_number, clone, treatment), as.character)) %>%
  dplyr::select(-Cage_ID) %>%
  pivot_longer(cols = -c(Animal_number, clone, treatment), names_to = "date", values_to = "weight") %>%
  mutate(
    date = as.Date(as.numeric(date), origin = "1899-12-30"),
    weight = as.numeric(weight)
  ) %>%
  mutate(day = as.numeric(date - min(date)) + 1)

df$clone <- factor(df$clone, levels=c("WT", "C3"))
df$treatment <- factor(df$treatment, levels=c("control", "treat"))

p <- ggplot(df, aes(x = day, y = weight, group = Animal_number, color = treatment)) +
  geom_line(alpha = 0.4, size = 0.6) +   # individual animals
  
  # SD ribbon
  stat_summary(
    aes(group = treatment, fill = treatment),
    fun.data = mean_sdl,
    fun.args = list(mult = 1),
    geom = "ribbon",
    alpha = 0.2,
    color = NA
  ) +
  
  # Mean line
  stat_summary(
    aes(group = treatment),
    fun = mean,
    geom = "line",
    linewidth = 1.5
  ) +
  
  theme_publication() +
  labs(title = "Animal Weight Over Time",
       x = "Days after Tumor Implantation",
       y = "Weight (g)",
       color = "Treatment",
       fill = "Treatment") +
  scale_color_manual(values = c("#377EB8","#E41A1C")) +
  facet_wrap(~clone)

ggsave("20260205_invivo_expt/animal_weight.pdf",
       plot = p, 
       width = 10,
       height = 10,
       dpi = 300,
       units = "cm")  

```


## Kras IF
```{r}
dat_pm <- read.csv("../../Data/KRAS_IF/output_dist/Intensity_CellmembraneArea.csv")
dat_pm$subcell <- "Plasma membrane"
dat_cyt <- read.csv("../../Data/KRAS_IF/output_dist/Intensity_CytoplasmArea.csv")
dat_cyt$subcell <- "Cytoplasm"



dat <- dat_pm %>%
  bind_rows(dat_cyt) %>%
  group_by(Metadata_Clone, Metadata_Cell_Nr, subcell) %>%
  summarise(mean_value = median(Intensity_MedianIntensity_Kras)) %>%
  ungroup()

dat$Metadata_Clone <- factor(dat$Metadata_Clone, levels=c("WT", "C3", "C5","C6"))
dat$group <- ifelse(dat$Metadata_Clone %in% "WT", "WT","KO")
dat$group <- factor(dat$group, levels=c("WT", "KO"))

p <- ggplot(data=dat) +
  geom_boxplot(aes(x= group, 
                  y= mean_value,
                  fill=group), 
               color="black",
               show.legend = FALSE) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  ggnewscale::new_scale_fill() +
  geom_jitter(aes(x= group, 
                  y= mean_value, 
                  fill=Metadata_Clone),
              size=3, 
              width = 0.4, 
              shape = 21, 
              color= "black",
              alpha= 0.8) +
  stat_compare_means(data=dat, 
                     aes(x= group, 
                         y= mean_value), 
                     label = "..p..",
                     show.legend = FALSE) +
  scale_fill_manual(values = c("#D1E5F0","#B2182B","#EF8A62","#FDDBC7")) +
  theme_publication() +
  xlab("") +
  labs(fill="")+
  ylab("Median fluorescence Intensity [AU]") +
  ggtitle("Kras accumulation") +
  facet_wrap(~subcell)


print(p)

ggsave("kras_cytoplasm.pdf",
       plot = p, 
       width = 9,
       height = 8,
       dpi = 300,
       units = "cm")
```

## Organoids treatment
```{r}
df <- read_excel("20260225_organoids/organoid_data.xlsx")

df$group <- factor(df$group, levels = c("Control", "MRTX1133",
                                        "ARC39", "ARC39+MRTX1133"))



# compute per-facet (POB) y-range so label offsets scale properly
ranges <- df %>%
  group_by(POB) %>%
  summarize(ymin = min(value, na.rm = TRUE),
            ymax = max(value, na.rm = TRUE),
            .groups = "drop")

# compute the group-median per (POB, group) and pick the lowest group-median per POB
hlines <- df %>%
  group_by(POB, group) %>%
  summarize(group_med = median(value, na.rm = TRUE), .groups = "drop") %>%
  group_by(POB) %>%
  slice_min(group_med, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  rename(median_lowest = group_med) %>%
  left_join(ranges, by = "POB") %>%
  mutate(offset = 0.05 * (ymax - ymin),   # 5% of y-range; tweak this if needed
         x = Inf)                         # place labels near the right edge of each facet

# original plot (your code)
p <- ggplot(data = df) +
  geom_boxplot(aes(x = group, y = value, fill = group),
               color = "black", show.legend = FALSE) +
  geom_jitter(aes(x = group, y = value, fill = group),
              size = 3, width = 0.4, shape = 21, color = "black", alpha = 0.8,
              show.legend = FALSE) +
  scale_fill_manual(values = c("#E41A1C","#3182BD", "#9ECAE1" ,"#DEEBF7" )) +
  stat_compare_means(data = df,
                     aes(x = group, y = value),
                     label = "..p..",
                     ref.group = "Control",
                     method = "t.test",
                     angle = 90,
                     show.legend = FALSE) +
  theme_publication() +    
  xlab("") +
  labs(fill = "") +
  ylab("% Cell viability") +
  facet_wrap(~POB)

# add horizontal line and labels per facet
p2 <- p +
  geom_hline(data = hlines,
             aes(yintercept = median_lowest),
            color = "black",
             inherit.aes = FALSE) +
  # label above the line
  geom_text(data = hlines,
            aes(x = x, y = median_lowest + offset, label = "antagonism"),
            hjust = 1.1, vjust = 0, size = 3.5, inherit.aes = FALSE) +
  # label below the line
  geom_text(data = hlines,
            aes(x = x, y = median_lowest - offset, label = "synergism"),
            hjust = 1.1, vjust = 1, size = 3.5, inherit.aes = FALSE)

print(p2)


ggsave("20260225_organoids/treatment.pdf",
       plot = p2, 
       width = 12,
       height = 9,
       dpi = 300,
       units = "cm")
```

## Classical Cells EMT
```{r}

library(ggpubr)
library(ggsignif)

df <- read.csv("20260228_EMT_SM_Ceramide/output/Intensity_Cells.csv") %>%
  dplyr::select(Metadata_cell_line, Metadata_treatment,
                Intensity_MeanIntensity_CER,Intensity_MeanIntensity_SM ) %>%
  pivot_longer(cols=c("Intensity_MeanIntensity_CER","Intensity_MeanIntensity_SM"), values_to = "intensity", names_to = "expression") %>%
  mutate(expression = gsub("Intensity_MeanIntensity_","", expression))



p <- ggplot(df,
            aes(x=expression,
                y=intensity,
                fill=Metadata_treatment)) +
  geom_boxplot()+
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  theme_publication() +
  xlab("") +
  labs(fill="")+
  ylab("Mean fluorescence Intensity [AU]") +
  facet_wrap(~Metadata_cell_line) 

stat.test <- df %>%
  group_by(expression, Metadata_cell_line) %>%
  t_test(intensity ~ Metadata_treatment) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj") %>%
  add_xy_position(x = "expression", dodge = 0.8)


p <- p + stat_pvalue_manual(
  stat.test,  label = "p", tip.length = 0
  ) 

ggsave("20260228_EMT_SM_Ceramide/cytoplasm.pdf",
       plot = p, 
       width = 10,
       height = 8,
       dpi = 300,
       units = "cm")
```

## SEM analysis
```{r}
df <- read_excel("20251219_TEM/analysis of SEM.xlsx") %>%
  separate(`Cell ID`, into = c("clone", "image", "replicate")) %>%
  mutate(group = ifelse(clone == "WT", "WT", "KO"))

df$group <- factor(df$group, c("WT", "KO"))


p <- ggplot(data=df) +
  geom_boxplot(aes(x= group, 
                  y= `outward invasion in %`,
                  fill=group), 
               color="black",
               show.legend = FALSE) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  ggnewscale::new_scale_fill() +
  geom_jitter(aes(x= group, 
                  y= `outward invasion in %`, 
                  fill=clone),
              #size=3, 
              width = 0.4, 
              shape = 21, 
              color= "black",
              alpha= 0.8) +
  stat_compare_means(data=df, 
                     aes(x= group, 
                         y= `outward invasion in %`), 
                     label = "..p..",
                     show.legend = FALSE) +
  scale_fill_manual(values = c("#B2182B","#EF8A62","#FDDBC7", "#D1E5F0")) +
  theme_publication() +
  xlab("")+
  ylab("% of Protrusions per um2") +
  ggtitle("Cell Protrusions") 

print(p)

ggsave("20251219_TEM/cell_protrusion.pdf",
       plot = p, 
       width = 9,
       height = 10,
       dpi = 300,
       units = "cm") 

```

## SMPD1 stage dependent 
```{r}
metadata <- readRDS("../../Data/TCGA/parsed_data_mauer_phenodata.rds")
dat <- readRDS("../../Data/TCGA/parsed_data_maurer_normalized.rds") %>%
  as.data.frame() 

dat <- log2(dat+1)

dat <- dat["SMPD1",]

library(dplyr)

dat <- dat %>%
  sjmisc::rotate_df() %>%
  rownames_to_column("Sample.Name") %>%
  inner_join(metadata, by = "Sample.Name") %>%
  filter(compartment == "Epithelium") %>%
  drop_na(AJCC_Stage) %>%
  mutate(
    stage = case_when(
      AJCC_Stage %in% c("1A","2A", "2B")    ~ "I-II",
      AJCC_Stage %in% c("3", "4")  ~ ">II",
      TRUE ~ NA_character_)
  )


dat$stage <- factor (dat$stage, levels = c("I-II", ">II"))


 p <- ggplot(dat,
             aes(x=stage,
                 y=SMPD1, 
                 fill = stage)) +
   geom_boxplot() +
   geom_jitter(shape=21,
               color="black", alpha=0.5) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  stat_compare_means(data=dat, 
                     aes(x= stage, 
                         y= SMPD1), 
                     label = "..p..",
                     show.legend = FALSE) +
  theme_publication() +
  xlab("") +
  labs(fill="")+
  ylab("log2 Normalized expression") +
  ggtitle("SMPD1") 
   
print(p)

ggsave("SMPD1_expression_Mauer.pdf",
       plot = p, 
       width = 10,
       height = 8,
       dpi = 300,
       units = "cm") 
```

## ARC39 proliferation

```{r}
df <- readxl::read_xlsx("20260227_MTT_activity/MTT-summary-ARC-39.xlsx", 3)



### replace zero to 0.001 for concentration
df$Dose[df$Dose == 0] <-0.001




### mean and sd
mtt.dat <- df %>%
  filter(Hours == 96) %>%
  filter(Group == "WT") %>%
  group_by(Group, Dose, Hours) %>%
  summarize(mean = mean(MTT),
            sd = sd(MTT))




p <- ggplot(mtt.dat, 
            aes(
              x = Dose, 
              y = mean, 
              color = Group
    )) +
      geom_errorbar(
        aes(ymax = mean + sd, ymin = mean - sd),
        lwd = 1,
        alpha = 0.5
      ) +
      scale_color_manual(values = c("#377EB8","#E41A1C")) +
      geom_jitter(
        data = mtt.dat,
        aes(x = Dose, y = mean, fill = Group),
        size = 2,
        width = 0.1,
        shape = 21,
        color = "black"
      ) +
   scale_fill_manual(values = c("#377EB8","#E41A1C")) +
      stat_compare_means(
        data = mtt.dat,
        aes(x = Dose,
            y = mean, group = Group),
        label = "p.signif",
        method = "t.test"
      ) +
      theme_publication() +
      xlab("concentration (uM) ") +
      ylab("% MTT Proliferation") +
      scale_x_continuous(
        breaks = log_breaks(n = 10),
        trans = log10_trans(),
        labels = function(n) {
          format(n, scientific = FALSE)
        }
      ) +
      scale_y_continuous(expand = c(0, 0), limits = c(0, 120)) +
      stat_smooth(
        method = "drm",
        method.args = list(fct = L.4(
          names = c("hill", "min_value", "max_value", "ec_50")
        )),
        se = FALSE,
        aes(y = mean, color = Group),
        size = 2
      ) +
      geom_hline(yintercept = 50, lty = "dotted") +
      annotation_logticks(sides = "b") +
      ggtitle("Proliferation ARC39") +
      theme(legend.position = c(0.1, 0.2))

  print(p)
  
  ggsave("20260227_MTT_activity/mtt_proliferation.pdf",
       plot = p, 
       width = 10,
       height = 8,
       dpi = 300,
       units = "cm")
  
  
  
  ED50 <-
      drm(mean ~ Dose,
          Group,
          data = mtt.dat,
          fct = LL.4(names = c("b", "lower", "upper", "ed50")))
    
summary(ED50)
    
    ed.summary <- ED(ED50, c(50), interval = "delta")
    
    

```




## ARC39 Activity

```{r}
df <- readxl::read_xlsx("20260227_MTT_activity/Smpd1 activity-ARC39-analysis.xlsx", 2)



### replace zero to 0.001 for concentration
df$Dose[df$Dose == 0] <-0.001




### mean and sd
mtt.dat <- df %>%
  filter(Group == "WT") %>%
  group_by(Group, Dose) %>%
  summarize(mean = mean(Activity),
            sd = sd(Activity))




p <- ggplot(mtt.dat, 
            aes(
              x = Dose, 
              y = mean, 
              color = Group
    )) +
      geom_errorbar(
        aes(ymax = mean + sd, ymin = mean - sd),
        lwd = 1,
        alpha = 0.5
      ) +
      scale_color_manual(values = c("#E41A1C")) +
      geom_jitter(
        data = mtt.dat,
        aes(x = Dose, y = mean, fill = Group),
        size = 2,
        width = 0.1,
        shape = 21,
        color = "black"
      ) +
   scale_fill_manual(values = c("#E41A1C")) +
      stat_compare_means(
        data = mtt.dat,
        aes(x = Dose,
            y = mean, group = Group),
        label = "p.signif",
        method = "t.test"
      ) +
      theme_publication() +
      xlab("concentration (uM) ") +
      ylab("% Smpd1 Activity") +
      scale_x_continuous(
        breaks = log_breaks(n = 10),
        trans = log10_trans(),
        labels = function(n) {
          format(n, scientific = FALSE)
        }
      ) +
      scale_y_continuous(expand = c(0, 0), limits = c(0, 120)) +
      stat_smooth(
        method = "drm",
        method.args = list(fct = L.4(
          names = c("hill", "min_value", "max_value", "ec_50")
        )),
        se = FALSE,
        aes(y = mean, color = Group),
        size = 2
      ) +
      geom_hline(yintercept = 50, lty = "dotted") +
      annotation_logticks(sides = "b") +
      ggtitle("ARC39 Activity") +
      theme(legend.position = c(0.1, 0.2))

  print(p)
  
  ggsave("20260227_MTT_activity/arc_smpd1_activity.pdf",
       plot = p, 
       width = 10,
       height = 8,
       dpi = 300,
       units = "cm")
  
  
  
  ED50 <-
      drm(mean ~ Dose,
          Group,
          data = mtt.dat,
          fct = LL.4(names = c("b", "lower", "upper", "ed50")))
    
summary(ED50)
    
    ed.summary <- ED(ED50, c(50), interval = "delta")
    
    

```

## human metabolome
```{r}

cells.dat <- read_xlsx("../../Data/Metabolome/human_cell_lines/KLUM-01-19MDCLP CLP CDT.XLSX", 
                        "Species Concentrations", 
                        skip=20)

cells.metadata <- read_xlsx("../../Data/Metabolome/human_cell_lines/KLUM-01-19MDCLP HD4 CDT.XLSX",  "metadata")

imputed.data<- cells.dat %>%
  dplyr::select(-Class, - `Group Units`) %>% 
  mutate(across(where(is.numeric), ~replace_na(.,0))) 

colnames(imputed.data)[1] <- "MetaboliteName"

metadata <- cells.metadata 
## assign column names
#colnames(imputed.data) <- metadata$CELL_GENOTYPE

library(metapacR)
## annotate chemical metadata
data("chemicalMetadata")
metabolite.class <- force(chemicalMetadata)

metabolite.class <- metabolite.class %>%
  dplyr::mutate(across(everything(), as.character)) %>%
  dplyr::rename(c("Metabolite" = "MET_CHEM_NO",
           "MetaboliteName" = "CHEMICAL_NAME")) %>%
  mutate_all(na_if,"") %>%
  distinct()

imputed.data <- imputed.data %>%
  inner_join(metabolite.class, by = "MetaboliteName")

##----------------------------------------------------------------
##                      total plot                      --
##----------------------------------------------------------------
imputed.data.total <- imputed.data %>%
  filter(SUPER_PATHWAY %in% c("Lipid", "Complex lipids")) %>%
  filter(grepl("^CER|^SM",MetaboliteName)) %>%
  separate(col=MetaboliteName, into = c("Metabolite", "chain_length"), sep = "[(]") %>%
  dplyr::mutate(chain_length = gsub("[)]", "", chain_length))  %>%
  filter(Metabolite %in% c("CER", "SM")) %>%
  distinct() 


df <- imputed.data.total[,1:43] %>%
  pivot_longer(cols= -c("Metabolite", "chain_length"), names_to = "CLIENT IDENTIFIER", values_to = "values") %>%
  inner_join(metadata, by = "CLIENT IDENTIFIER") %>%
  filter(`CELL LINE` %in% c("PaTuT", "PaTuS")) %>%
  filter(KO_STATUS %in% c("WT")) %>%
  dplyr::select(Metabolite, chain_length,`CLIENT IDENTIFIER`,`CELL LINE`,chain_length,values) %>%
  distinct()

df <- janitor::clean_names(df)

# 1) Mean ± SD per diagnosis (2 decimals)
stats_wide <- df %>%
  group_by(metabolite, chain_length, cell_line) %>%
  summarise(
    mean = mean(values, na.rm = TRUE),
    sd   = sd(values, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(`Mean ± SD` = sprintf("%.2f ± %.2f", mean, sd)) %>%
  dplyr::select(metabolite, chain_length, cell_line, `Mean ± SD`) %>%
  pivot_wider(names_from = cell_line, values_from = `Mean ± SD`)

# 2) P-values (3 decimals, with <0.001 formatting)
pvals <- df %>%
  filter(!is.na(values), !is.na(cell_line)) %>%
  group_by(metabolite, chain_length) %>%
  summarise(
    p_raw = {
      dx <- droplevels(factor(cell_line))
      if (nlevels(dx) != 2) NA_real_ else
        tryCatch(t.test(values ~ dx)$p.value, error = function(e) NA_real_)
    },
    .groups = "drop"
  ) %>%
  mutate(
    p_adj = p.adjust(p_raw, method = "BH"),
    # Formatting function for 3 decimals or <0.001
    p_value = sprintf("%.4f", p_raw),
    p_fdr   = sprintf("%.4f", p_adj)
  ) %>%
  dplyr::select(metabolite, chain_length, p_value, p_fdr)

# 3) Final summary table
summary_table <- stats_wide %>%
  left_join(pvals, by = c("metabolite", "chain_length")) %>%
  arrange(metabolite, chain_length)

write.csv(summary_table, "summary_table_sm_cer_cell_line.csv")
  

df$cell_line <- factor(df$cell_line, levels = c("PaTuT", "PaTuS"))

p <- ggplot(df,
            aes(x= chain_length,
                y=values,
                fill=cell_line)) +
  geom_boxplot(size=0.1) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  theme_publication() +
  xlab("") +
  labs(fill="")+
  ylab("Concentration in nmol/mg of protein")

stat.test <- df %>%
  group_by(chain_length, metabolite) %>%
  t_test(values ~ cell_line) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj") %>%
  add_xy_position(x = "chain_length", dodge = 0.8) +


p <- p + stat_pvalue_manual(
  stat.test,  label = "p", tip.length = 0,
  angle = 90
  )  +
  facet_grid(~metabolite, scales="free")

print(p)

ggsave("cer_sm_individual_abs.pdf",
       plot = p, 
       width = 15,
       height = 8,
       dpi = 300,
       units = "cm")


### ratio


df <- df %>%
  group_by(chain_length, client_identifier, cell_line) %>%
  summarise_if(is.numeric,funs(.[metabolite == "CER"] / .[metabolite == "SM"])) %>%
  ungroup()
  
df[sapply(df, is.infinite)] <- NA 


df$cell_line <- factor(df$cell_line, levels = c("PaTuT", "PaTuS"))


# 1) Mean ± SD per diagnosis (2 decimals)
stats_wide <- df %>%
  group_by(chain_length, cell_line) %>%
  summarise(
    mean = mean(values, na.rm = TRUE),
    sd   = sd(values, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(`Mean ± SD` = sprintf("%.2f ± %.2f", mean, sd)) %>%
  dplyr::select(chain_length, cell_line, `Mean ± SD`) %>%
  pivot_wider(names_from = cell_line, values_from = `Mean ± SD`)

# 2) P-values (3 decimals, with <0.001 formatting)
pvals <- df %>%
  filter(!is.na(values), !is.na(cell_line)) %>%
  group_by(chain_length) %>%
  summarise(
    p_raw = {
      dx <- droplevels(factor(cell_line))
      if (nlevels(dx) != 2) NA_real_ else
        tryCatch(t.test(values ~ dx)$p.value, error = function(e) NA_real_)
    },
    .groups = "drop"
  ) %>%
  mutate(
    p_adj = p.adjust(p_raw, method = "BH"),
    # Formatting function for 3 decimals or <0.001
    p_value = sprintf("%.4f", p_raw),
    p_fdr   = sprintf("%.4f", p_adj)
  ) %>%
  dplyr::select(chain_length, p_value, p_fdr)

# 3) Final summary table
summary_table <- stats_wide %>%
  left_join(pvals, by = c("chain_length")) %>%
  arrange(chain_length)

write.csv(summary_table, "summary_table_sm_cer_ratio_cell_line.csv")
  

p <- ggplot(df,
            aes(x= chain_length,
                y=values,
                fill=cell_line)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  theme_publication() +
  xlab("") +
  labs(fill="")+
  ylab("CER/SM ratio")

stat.test <- df %>%
  group_by(chain_length) %>%
  t_test(values ~ cell_line) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj") %>%
  add_xy_position(x = "chain_length", dodge = 0.8)


p <- p + stat_pvalue_manual(
  stat.test,  label = "p", tip.length = 0,
  angle = 90
  ) 

print(p)

ggsave("cer_sm_ratio_abs.pdf",
       plot = p, 
       width = 15,
       height = 8,
       dpi = 300,
       units = "cm")
```

## liposomes
```{r}
dat <- readxl::read_excel("20260228_liposomes/Liposom_Cermide_staining_Results.xlsx")


dat$`CER/SM ratio` <- factor (dat$`CER/SM ratio`, levels = c("50:50", "90:10"))


 p <- ggplot(dat,
             aes(x=`CER/SM ratio`,
                 y=`Average of Intensity_MeanIntensity_liposome`, 
                 fill = `CER/SM ratio`)) +
   geom_violin(alpha=0.5) +
   geom_boxplot(
                width=0.1,
               color="black") +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  stat_compare_means(data=dat, 
                     aes(x=`CER/SM ratio`,
                 y=`Average of Intensity_MeanIntensity_liposome`), 
                     label = "..p..",
                     show.legend = FALSE) +
  theme_publication() +
  xlab("") +
  labs(fill="")+
  ylab("Mean Flouroscence Intensity / liposome") +
  ggtitle("Ceramide IF") 
   
print(p)

ggsave("20260228_liposomes/ceramide_if.pdf",
       plot = p, 
       width = 10,
       height = 8,
       dpi = 300,
       units = "cm") 
```
## human metabolome METAPAC
```{r}

dat <- readRDS("Metabolon_data/expData_metabolon_untargeted.rds")

imputed.data<- dat$raw.counts %>%
  #mutate(across(where(is.numeric), ~replace_na(.,0))) %>%
  sjmisc::rotate_df(cn=TRUE, rn="MetaboliteName")

metadata <- dat$metadata

library(metapacR)
## annotate chemical metadata
data("chemicalMetadata")
metabolite.class <- force(chemicalMetadata)

metabolite.class <- metabolite.class %>%
  dplyr::mutate(across(everything(), as.character)) %>%
  dplyr::rename(c("Metabolite" = "MET_CHEM_NO",
           "MetaboliteName" = "CHEMICAL_NAME")) %>%
  mutate_all(na_if,"") %>%
  distinct()

imputed.data <- imputed.data %>%
  inner_join(metabolite.class, by = "MetaboliteName")

##----------------------------------------------------------------
##                      total plot                      --
##----------------------------------------------------------------
imputed.data.total <- imputed.data %>%
  filter(SUPER_PATHWAY %in% c("Lipid", "Complex lipids")) %>%
  filter(grepl("^CER|^SM",MetaboliteName)) %>%
  separate(col=MetaboliteName, into = c("Metabolite", "chain_length"), sep = "[(]") %>%
  dplyr::mutate(chain_length = gsub("[)]", "", chain_length))  %>%
  filter(Metabolite %in% c("CER", "SM")) %>%
  distinct() 


df <- imputed.data.total[,1:1327] %>%
  pivot_longer(cols= -c("Metabolite", "chain_length"), names_to = "LB01=LB03", values_to = "values") %>%
  inner_join(metadata, by = "LB01=LB03") %>%
  mutate(diagnosis=if_else(cummulative_diagnosis == "Pankreaskarzinom", "PDAC", "Rest")) %>%
  dplyr::select(Metabolite, chain_length,`LB01=LB03`,diagnosis,chain_length,values) %>%
  distinct() %>%
  drop_na(diagnosis) %>%
  ungroup()


# 1) Mean ± SD per diagnosis (2 decimals)
stats_wide <- df %>%
  group_by(Metabolite, chain_length, diagnosis) %>%
  summarise(
    mean = mean(values, na.rm = TRUE),
    sd   = sd(values, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(`Mean ± SD` = sprintf("%.2f ± %.2f", mean, sd)) %>%
  select(Metabolite, chain_length, diagnosis, `Mean ± SD`) %>%
  pivot_wider(names_from = diagnosis, values_from = `Mean ± SD`)

# 2) P-values (3 decimals, with <0.001 formatting)
pvals <- df %>%
  filter(!is.na(values), !is.na(diagnosis)) %>%
  group_by(Metabolite, chain_length) %>%
  summarise(
    p_raw = {
      dx <- droplevels(factor(diagnosis))
      if (nlevels(dx) != 2) NA_real_ else
        tryCatch(t.test(values ~ dx)$p.value, error = function(e) NA_real_)
    },
    .groups = "drop"
  ) %>%
  mutate(
    p_adj = p.adjust(p_raw, method = "BH"),
    # Formatting function for 3 decimals or <0.001
    p_value = sprintf("%.4f", p_raw),
    p_fdr   = sprintf("%.4f", p_adj)
  ) %>%
  select(Metabolite, chain_length, p_value, p_fdr)

# 3) Final summary table
summary_table <- stats_wide %>%
  left_join(pvals, by = c("Metabolite", "chain_length")) %>%
  arrange(Metabolite, chain_length)

write.csv(summary_table, "Metabolon_data/summary_table_sm_cer.csv")

df$diagnosis <- factor(df$diagnosis, levels = c("PDAC", "Rest"))

writexl::write_xlsx(df, "Metabolon_data/Cer_sm_qunatification.xlsx")

p <- ggplot(df,
            aes(x= chain_length,
                y=values,
                fill=diagnosis)) +
  geom_boxplot(size=0.1) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  theme_publication() +
  xlab("") +
  labs(fill="")+
  ylab("Concentration in uM") 

stat.test <- df %>%
  group_by(chain_length, Metabolite) %>%
  t_test(values ~ diagnosis) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj") %>%
  add_xy_position(x = "chain_length", dodge = 0.8)


p <- p   +
  facet_wrap(~Metabolite, scale="free", ncol = 1) + stat_pvalue_manual(
  stat.test,  label = "p", tip.length = 0,
  angle = 90
  )

print(p)

ggsave("Metabolon_data/cer_sm_individual_abs.pdf",
       plot = p, 
       width = 15,
       height = 15,
       dpi = 300,
       units = "cm")


### ratio

df <- df %>%
  group_by(chain_length, `LB01=LB03`, diagnosis) %>%
  summarise_if(is.numeric,funs(.[Metabolite == "CER"] / .[Metabolite == "SM"])) %>%
  ungroup()
  
df[sapply(df, is.infinite)] <- 0


df$diagnosis <- factor(df$diagnosis, levels = c("PDAC", "Rest"))


p <- ggplot(df,
            aes(x= chain_length,
                y=values,
                fill=diagnosis)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  theme_publication() +
  xlab("") +
  labs(fill="")+
  ylab("CER/SM ratio")

stat.test <- df %>%
  group_by(chain_length) %>%
  t_test(values ~ diagnosis) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj") %>%
  add_xy_position(x = "chain_length", dodge = 0.8)


p <- p + stat_pvalue_manual(
  stat.test,  label = "p", tip.length = 0,
  angle = 90
  ) + scale_y_continuous(
        breaks = log_breaks(n = 10),
        trans = log10_trans(),
        labels = function(n) {
          format(n, scientific = FALSE)
        }
      )

print(p)

ggsave("Metabolon_data/cer_sm_ratio_abs.pdf",
       plot = p, 
       width = 15,
       height = 8,
       dpi = 300,
       units = "cm")

```

## Proteome heatmap

```{r}
save_pheatmap_pdf <- function(x, filename, width=7, height=7) {
   stopifnot(!missing(x))
   stopifnot(!missing(filename))
   pdf(filename, width=width, height=height)
   grid::grid.newpage()
   grid::grid.draw(x$gtable)
   dev.off()
}

df <- readxl::read_excel("20260202_proteome/Proteins_HeatMap.xlsx", "All")

metadata <- readxl::read_excel("20260202_proteome/Proteins_HeatMap.xlsx", "metadata")

phenodata <- df[, 1:5] %>%
  column_to_rownames("T: Genes")


mat <- as.matrix(sapply(df[, 6:23], as.numeric))

mat[is.nan(mat)] <- 0

rownames(mat) <- rownames(phenodata)

ann <- data.frame(Status = metadata$Group,
                  Clone =metadata$Clone, 
                  Treatment=metadata$Treatment)

rownames(ann) <- colnames(mat)

ann_row <- data.frame(Cluster = phenodata$Catagories)
rownames(ann_row) <- rownames(phenodata)

Status.col <- c("#377EB8","#E41A1C")
names(Status.col) <- unique(ann$Status)


Treatment.col <- c("#66C2A5","#FC8D62")
names(Treatment.col) <- unique(ann$Treatment)

Clone.col <- c("#D1E5F0","#B2182B","#EF8A62","#FDDBC7")
names(Clone.col) <- unique(ann$Clone)

Cluster.col <- brewer.pal(9, "Set1")
names(Cluster.col) <- unique(ann_row$Cluster)


combo.cols <- list(Status = Status.col,
                     Clone = Clone.col,
                   Treatment = Treatment.col,
                   Cluster = Cluster.col)

p <- pheatmap::pheatmap(mat, 
                   annotation_col = ann, 
                   annotation_row = ann_row,
                   annotation_colors = combo.cols,
                   scale = "row", 
                   col = colorRampPalette(rev(brewer.pal(40, "RdBu")))(40),
                   fontsize_row = 6,
                   fontsize_col = 6,
                   border_color = NA,
                   cluster_rows = FALSE, 
                   cluster_cols = FALSE,
                   show_colnames = FALSE)
  
  print(p)
  
  save_pheatmap_pdf(p,
                  "20260202_proteome/heatmap.pdf",
                  width=5,
                  height=6)

```

## ihc staining
```{r}
colony <- read_excel("20260205_invivo_expt/aSMA-measurements-orthotop-analysis.xlsx")
colony$group <- factor(colony$group, levels=c("WT", "KO"))
colony$strain <- factor(colony$strain, levels=c("WT", "Smpd1"))

colony$value <- colony$`total aSMA positive in %`

colony <- colony %>%
  filter(!clone == "C3" | !strain =="Smpd1",)
 

p <- ggplot(data=colony) +
  geom_boxplot(aes(x= group, 
                  y= value,
                  fill=group), 
               color="black",
               show.legend = FALSE) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  ggnewscale::new_scale_fill() +
  geom_jitter(aes(x= group, 
                  y= value, 
                  fill=clone),
              size=3, 
              width = 0.4, 
              shape = 21, 
              color= "black",
              alpha= 0.8) +
  stat_compare_means(data=colony, 
                     aes(x= group, 
                         y= value), 
                     label = "..p..",
                     show.legend = FALSE) +
  scale_fill_manual(values = c("#B2182B", "#D1E5F0")) +
  theme_bw() +
  xlab("") +
  ylab("total aSMA positive in %") +
  ggtitle("aSMA") +
  facet_grid(~strain, ,
             scales="free_x",
             space='free_x')

print(p)

ggsave("20260205_invivo_expt/aSMA_positive.pdf",
       plot = p, 
       width = 10,
       height = 7,
       dpi = 300,
       units = "cm")

colony <- read_excel("20260205_invivo_expt/CD31-measurements-orthotop analysis.xlsx")
colony$group <- factor(colony$group, levels=c("WT", "KO"))
colony$strain <- factor(colony$strain, levels=c("WT", "Smpd1"))

colony$value <- colony$`total CD31 positive in %`

colony <- colony %>%
  filter(!clone == "C3" | !strain =="Smpd1",)

p <- ggplot(data=colony) +
  geom_boxplot(aes(x= group, 
                  y= value,
                  fill=group), 
               color="black",
               show.legend = FALSE) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  ggnewscale::new_scale_fill() +
  geom_jitter(aes(x= group, 
                  y= value, 
                  fill=clone),
              size=3, 
              width = 0.4, 
              shape = 21, 
              color= "black",
              alpha= 0.8) +
  stat_compare_means(data=colony, 
                     aes(x= group, 
                         y= value), 
                     label = "..p..",
                     show.legend = FALSE) +
  scale_fill_manual(values = c("#B2182B", "#D1E5F0")) +
  theme_bw() +
  xlab("") +
  ylab("total CD31 positive in %") +
  ggtitle("Cd31") +
  facet_grid(~strain, ,
             scales="free_x",
             space='free_x')

print(p)

ggsave("20260205_invivo_expt/CD31_positive.pdf",
       plot = p, 
       width = 10,
       height = 7,
       dpi = 300,
       units = "cm")
```

## MRTX resistanct cells

```{r}
scriptLibraries <-  c("here",
                      "tidyverse",
                      "GEOquery",
                      "biomaRt",
                      "SummarizedExperiment",
                      "DESeq2",
                      "limma")
##---------------------------------------------------------------
##                      load functions                    --
##---------------------------------------------------------------
source("https://raw.githubusercontent.com/umahajanatlmu/useful_commands/main/auxillary/basicFunctions.R")
##---------------------------------------------------------------
##                        load packages                        --
##---------------------------------------------------------------
installScriptLibs(scriptLibraries)
```


### human cells
```{r}
files <- list.files("20260228_MRTX_res/GSE269985_RAW/Human",
                    full.names = TRUE)
i <- 1
 
for (f in files) {
  gsm <- gsub(".*(GSM[0-9]+).*", "\\1", f)
  
  if ( i == 1) {
    df <- read.csv(f) 
  
  colnames(df)[3] <- gsm
  } else {
    df1 <- read.csv(f) 
  
  colnames(df1)[3] <- gsm
  
  df <- df %>%
    inner_join(df1, by =c("gene", "ensembl_id"))
  }
  
  i <- i +1
}

df_h <- df %>%
  filter(if_all(everything(), ~ .x != "")) %>%
  dplyr::select(-ensembl_id) %>%
  distinct(gene, .keep_all = TRUE) %>%
  column_to_rownames("gene")

metadata <- read.csv("20260228_MRTX_res/SraRunTable.csv", sep=";")
metadata <- metadata[metadata$Sample.Name %in% colnames(df_h),]

df_h <- log2(df_h+1) %>%
  sjmisc::rotate_df() %>%
  dplyr::select("SMPD1") %>%
  rownames_to_column("Sample.Name") %>%
  inner_join(metadata,by ="Sample.Name") %>%
  mutate(organism= "human")

p <- ggplot(df_h,
            aes(x=cell_line,
                y=SMPD1,
                fill=Treatment)) +
  geom_boxplot(show.legend = FALSE) +
  geom_jitter(size=3, 
               position=position_jitterdodge(jitter.width = 0.5), 
              shape = 21, 
              color= "black",
              alpha= 0.5) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  stat_compare_means(data=df_h, 
                     aes(x= cell_line, 
                         y= SMPD1), 
                     label = "..p..",
                     show.legend = FALSE) +
  theme_publication() +
  xlab("") +
  ylab("SMPD1 Normalized Expression") +
  ggtitle("MRTx1133 resistance")

print(p)

ggsave("20260228_MRTX_res/resistance_smpd1.pdf",
       plot = p, 
       width = 9,
       height = 10,
       dpi = 300,
       units = "cm")
```


```{r}
files <- list.files("20260228_MRTX_res/GSE269985_RAW/Human",
                    full.names = TRUE)
i <- 1
 
for (f in files) {
  gsm <- gsub(".*(GSM[0-9]+).*", "\\1", f)
  
  if ( i == 1) {
    df <- read.csv(f) 
  
  colnames(df)[3] <- gsm
  } else {
    df1 <- read.csv(f) 
  
  colnames(df1)[3] <- gsm
  
  df <- df %>%
    inner_join(df1, by =c("gene", "ensembl_id"))
  }
  
  i <- i +1
}

df_h <- df %>%
  filter(if_all(everything(), ~ .x != "")) %>%
  dplyr::select(-ensembl_id) %>%
  distinct(gene, .keep_all = TRUE) %>%
  column_to_rownames("gene")

metadata <- read.csv("20260228_MRTX_res/SraRunTable.csv", sep=";")
metadata <- metadata[metadata$Sample.Name %in% colnames(df_h),]

df_h <- log2(df_h+1) %>%
  sjmisc::rotate_df() %>%
  dplyr::select("SMPD1") %>%
  rownames_to_column("Sample.Name") %>%
  inner_join(metadata,by ="Sample.Name") %>%
  mutate(organism= "human")

p <- ggplot(df_h,
            aes(x=cell_line,
                y=SMPD1,
                fill=Treatment)) +
  geom_boxplot(show.legend = FALSE) +
  geom_jitter(size=3, 
               position=position_jitterdodge(jitter.width = 0.5), 
              shape = 21, 
              color= "black",
              alpha= 0.5) +
  scale_fill_manual(values = c("#377EB8","#E41A1C")) +
  stat_compare_means(data=df_h, 
                     aes(x= cell_line, 
                         y= SMPD1), 
                     label = "..p..",
                     show.legend = FALSE) +
  theme_publication() +
  xlab("") +
  ylab("SMPD1 Normalized Expression") +
  ggtitle("MRTx1133 resistance")

print(p)

ggsave("20260228_MRTX_res/resistance_smpd1.pdf",
       plot = p, 
       width = 9,
       height = 10,
       dpi = 300,
       units = "cm")
```


## Akt Inh- WT vs KO

```{r}
df <- readxl::read_xlsx("20260226_Akt/summary-Akt-inhibitor-Jianying.xlsx", 3)



### replace zero to 0.001 for concentration
df$Dose[df$Dose == 0] <-0.001


### mean and sd
mtt.dat <- df %>%
  filter(Hours == 72) %>%
  group_by(Group, Dose, Hours) %>%
  summarize(mean = mean(MTT),
            sd = sd(MTT))


mtt.dat$Group <- factor(mtt.dat$Group, levels=c("WT", "KO"))

p <- ggplot(mtt.dat, 
            aes(
              x = Dose, 
              y = mean, 
              color = Group
    )) +
      geom_errorbar(
        aes(ymax = mean + sd, ymin = mean - sd),
        lwd = 1,
        alpha = 0.5
      ) +
      scale_color_manual(values = c("#377EB8","#E41A1C")) +
      geom_jitter(
        data = mtt.dat,
        aes(x = Dose, y = mean, fill = Group),
        size = 2,
        width = 0.1,
        shape = 21,
        color = "black"
      ) +
   scale_fill_manual(values = c("#377EB8","#E41A1C")) +
      stat_compare_means(
        data = mtt.dat,
        aes(x = Dose,
            y = mean, group = Group),
        label = "p.signif",
        method = "t.test"
      ) +
      theme_publication() +
      xlab("concentration (uM) ") +
      ylab("% MTT Proliferation") +
      scale_x_continuous(
        breaks = log_breaks(n = 10),
        trans = log10_trans(),
        labels = function(n) {
          format(n, scientific = FALSE)
        }
      ) +
      scale_y_continuous(expand = c(0, 0), limits = c(0, 120)) +
      stat_smooth(
        method = "drm",
        method.args = list(fct = L.4(
          names = c("hill", "min_value", "max_value", "ec_50")
        )),
        se = FALSE,
        aes(y = mean, color = Group),
        size = 2
      ) +
      geom_hline(yintercept = 50, lty = "dotted") +
      annotation_logticks(sides = "b") +
      ggtitle("Proliferation Akt Inh") +
      theme(legend.position = c(0.1, 0.2))

  print(p)
  
  ggsave("20260226_Akt/mtt_proliferation_KO.pdf",
       plot = p, 
       width = 10,
       height = 8,
       dpi = 300,
       units = "cm")
  
  
  
  ED50 <-
      drm(mean ~ Dose,
          Group,
          data = mtt.dat,
          fct = LL.4(names = c("b", "lower", "upper", "ed50")))
    
summary(ED50)
    
    ed.summary <- ED(ED50, c(50), interval = "delta")
    
    

```

## Synergism
```{r}
library(synergyfinder)
screening_data <- readxl::read_excel("20260226_Akt/ARC39-AKTinh-combination-analysis.xlsx", "rplot")

screening_data <- screening_data %>%
  filter(Hours == 72) %>%
  group_by(conc_r, conc_c) %>%
  mutate(response = median(response)) %>%
  ungroup() %>%
  unique()


res <- ReshapeData(
  data = screening_data,
  data_type = "viability",
  impute = TRUE,
  impute_method = NULL,
  noise = TRUE,
  seed = 1)

res <- CalculateSynergy(
  data = res,
  method = c("ZIP", "HSA", "Bliss", "Loewe"),
  Emin = NA,
  Emax = NA,
  correct_baseline = "non")

res1 <- CalculateSensitivity(
  data = res,
  correct_baseline = "non"
)

res1$drug_pairs

p <- Plot2DrugContour(
  data = res,
  plot_block = 1,
  drugs = c(1, 2),
  plot_value = "ZIP_synergy",
  dynamic = FALSE,
  summary_statistic = c("mean","quantile_25", "quantile_75")
) + theme_publication()  +
  geom_contour(colour = "black",
               size = 0.5) +
  scale_fill_gradientn(
    colours = colorRampPalette(rev(RColorBrewer::brewer.pal(10, "RdBu")))(100),
    oob = scales::squish,
    limits = c(-50, 50),
    name = "Synergy score"
  ) +
  guides(fill = guide_colourbar(
    barwidth = unit(0.3, "cm"),
    ticks.colour = "black",
    frame.colour = "black"
  ))


print(p)

ggsave("20260226_Akt/synergy_score_72.pdf",
       plot = p,
       width = 10, 
       height = 8,
       units = "cm",
       dpi = 300)


screening_data <- readxl::read_excel("20260226_Akt/ARC39-AKTinh-combination-analysis.xlsx", "rplot")

screening_data <- screening_data %>%
  filter(Hours == 96) %>%
  group_by(conc_r, conc_c) %>%
  mutate(response = median(response)) %>%
  ungroup() %>%
  unique()


res <- ReshapeData(
  data = screening_data,
  data_type = "viability",
  impute = TRUE,
  impute_method = NULL,
  noise = TRUE,
  seed = 1)

res <- CalculateSynergy(
  data = res,
  method = c("ZIP", "HSA", "Bliss", "Loewe"),
  Emin = NA,
  Emax = NA,
  correct_baseline = "non")

res1 <- CalculateSensitivity(
  data = res,
  correct_baseline = "non"
)

res1$drug_pairs

p <- Plot2DrugContour(
  data = res,
  plot_block = 1,
  drugs = c(1, 2),
  plot_value = "ZIP_synergy",
  dynamic = FALSE,
  summary_statistic = c("mean","quantile_25", "quantile_75")
) + theme_publication()  +
  geom_contour(colour = "black",
               size = 0.5) +
  scale_fill_gradientn(
    colours = colorRampPalette(rev(RColorBrewer::brewer.pal(10, "RdBu")))(100),
    oob = scales::squish,
    limits = c(-50, 50),
    name = "Synergy score"
  ) +
  guides(fill = guide_colourbar(
    barwidth = unit(0.3, "cm"),
    ticks.colour = "black",
    frame.colour = "black"
  ))


print(p)

ggsave("20260226_Akt/synergy_score_96.pdf",
       plot = p,
       width = 10, 
       height = 8,
       units = "cm",
       dpi = 300)

```

## qPCR
```{r}
df <- read_excel("20260228_qPCR/EMT-qpcr-results-calculation-20260227-Ahmed.xlsx", "rplot")

# 1. Calculate stats comparing only Group 2 and Group 3
# Replace "Group 2" and "Group 3" with the actual names in your 'group' column
stat.test <- df %>%
  filter(group %in% c("StemXvivo", "StemXvivo+ARC39")) %>%
  group_by(cell_line, gene) %>%
  t_test(ddct ~ group) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_xy_position(x = "cell_line", dodge = 0.1) 

# 2. Create the plot
p <- ggplot(df, aes(x = cell_line, y = ddct, fill = group)) +
  # Bar layer (all 3 groups will show because we use 'df' here)
  stat_summary(
    fun = mean, geom = "bar",
    position = position_dodge(0.9), 
    color = "black"
  ) +
  # Error bar layer
  stat_summary(
    fun.data = mean_sd, 
    geom = "errorbar",
    width = 0.25, 
    position = position_dodge(0.9)
  ) +
  # Add the p-values (only for the Group 2 vs Group 3 comparison)
  stat_pvalue_manual(
    stat.test, 
    label = "p"
  ) +
  scale_fill_manual(values = c("#377EB8", "#E41A1C", "#EF8A62")) +
  facet_wrap(~gene) +
  scale_y_continuous(trans = log2_trans()) +
  theme_publication() +
  labs(x = "", y = "Relative fold Changes", title = "EMT", fill = "")

print(p)


ggsave("20260228_qPCR/emt.pdf",
       plot = p, 
       width = 20,
       height = 12,
       dpi = 300,
       units = "cm") 

```



## Smpd1 Activity Resistance
```{r}
df <- read_excel("20260228_Smpd1_activities/Resistance cells-Smpd1 activity-analysis.xlsx", "rplot")

p <- ggplot(df) +
  geom_boxplot(aes(x=group,
                 y=rel_value, 
                 fill = group),
               color = "black") +
  scale_fill_manual(values = c("#377EB8","#E41A1C", "#EF8A62" )) +
  stat_compare_means(data=df, 
                     aes(x=group,
                         y=rel_value), 
                     label = "..p..",
                     show.legend = FALSE) +
  theme_publication() +
  xlab("") +
  labs(fill="")+
  ylab("Relative Changes") +
  ggtitle("SMPD1 activity in Resistant cells") 
   
print(p)

ggsave("20260228_Smpd1_activities/resistant_cells.pdf",
       plot = p, 
       width = 9,
       height = 10,
       dpi = 300,
       units = "cm") 

```




## Smpd1 Activity StemXVivo
```{r}
df <- read_excel("20260228_Smpd1_activities/PS+C1-B3.xlsx", "rplot")

p <- ggplot(df) +
  geom_boxplot(aes(x=cell_line,
                 y=relative_change, 
                 fill = group),
               color = "black") +
  scale_fill_manual(values = c("#377EB8","#E41A1C", "#EF8A62" )) +
  stat_compare_means(data=df, 
                     aes(x=cell_line,
                         y=relative_change), 
                     label = "..p..",
                     show.legend = FALSE) +
  theme_publication() +
  xlab("") +
  labs(fill="")+
  ylab("Relative Changes") +
  ggtitle("SMPD1 activity in Resistant cells") 
   
print(p)

ggsave("20260228_Smpd1_activities/stemXvivo.pdf",
       plot = p, 
       width = 10,
       height = 8,
       dpi = 300,
       units = "cm") 

```





## Smpd1 Sensitivity WT and KO
```{r}
df <- read_excel("20260228_Smpd1_activities/ARC39 treatment in WT and KO.xlsx", "rplot")

### replace zero to 0.001 for concentration
df$Dose[df$Dose == 0] <-0.001



### mean and sd
mtt.dat <- df %>%
  filter(Group == "KO") %>%
  group_by(Group, Dose) %>%
  summarize(mean = mean(Activity),
            sd = sd(Activity))

df$Group <- factor(df$Group, levels = c("WT", "KO"))

p <- ggplot(mtt.dat, 
            aes(
              x = Dose, 
              y = mean, 
              color = Group
    )) +
      geom_errorbar(
        aes(ymax = mean + sd, ymin = mean - sd),
        lwd = 1,
        alpha = 0.5
      ) +
      scale_color_manual(values = c("#E41A1C")) +
      geom_jitter(
        data = mtt.dat,
        aes(x = Dose, y = mean, fill = Group),
        size = 2,
        width = 0.1,
        shape = 21,
        color = "black"
      ) +
   scale_fill_manual(values = c("#E41A1C")) +
      stat_compare_means(
        data = mtt.dat,
        aes(x = Dose,
            y = mean, group = Group),
        label = "p.signif",
        method = "t.test"
      ) +
      theme_publication() +
      xlab("concentration (uM) ") +
      ylab("% Cell Viability") +
      scale_x_continuous(
        breaks = log_breaks(n = 10),
        trans = log10_trans(),
        labels = function(n) {
          format(n, scientific = FALSE)
        }
      ) +
      scale_y_continuous(expand = c(0, 0), limits = c(0, 120)) +
      stat_smooth(
        method = "drm",
        method.args = list(fct = L.4(
          names = c("hill", "min_value", "max_value", "ec_50")
        )),
        se = FALSE,
        aes(y = mean, color = Group),
        size = 2
      ) +
      geom_hline(yintercept = 50, lty = "dotted") +
      annotation_logticks(sides = "b") +
      ggtitle("Smpd1 Sensitivity") +
      theme(legend.position = c(0.1, 0.2))

  print(p)
  
  ggsave("20260228_Smpd1_activities/arc_wt_ko.pdf",
       plot = p, 
       width = 10,
       height = 8,
       dpi = 300,
       units = "cm")
  
  
  
  ED50 <-
      drm(mean ~ Dose,
          Group,
          data = mtt.dat,
          fct = LL.4(names = c("b", "lower", "upper", "ed50")))
    
summary(ED50)
    
    ed.summary <- ED(ED50, c(50), interval = "delta")
```


## PanIN quant
```{r}
df <- read_excel("20260205_invivo_expt/PanIN_Quant_Reviewer_only.xlsx")

# 1. Calculate stats comparing only Group 2 and Group 3
# Replace "Group 2" and "Group 3" with the actual names in your 'group' column
stat.test <- df %>%
  group_by(Timepoint) %>%
  t_test(`PanIN%` ~ Genotype) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_xy_position(x = "Timepoint", dodge = 0.1) 

df$Timepoint <- factor(df$Timepoint, levels = c(6,12))

# 2. Create the plot
p <- ggplot(df, aes(x = Timepoint, y = `PanIN%`, fill = Genotype)) +
  geom_boxplot() +
  # +
  # Add the p-values (only for the Group 2 vs Group 3 comparison)
  stat_pvalue_manual(
    stat.test, 
    label = "p"
  ) +
  scale_fill_manual(values = c("#377EB8", "#E41A1C")) +
  theme_publication() +
  labs(x = "", y = "% of PanIN/ mm2", title = "PanINs", fill = "")

print(p)


ggsave("20260205_invivo_expt/panin.pdf",
       plot = p, 
       width = 9,
       height = 9,
       dpi = 300,
       units = "cm") 

```





# computing environment

```{r}
sessionInfo()
```
